---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Load the data (i.e. read.csv())
```{r echo=TRUE}
#fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(fileUrl, "DataActivity.zip", mode="wb")
#unzip("DataActivity.zip")
activitywithNA <- read.csv("activity.csv")
```
2. Process/transform the data (if necessary) into a format suitable for analysis
```{r echo=TRUE}
# remove the records which does not have any data
activity <- activitywithNA[!is.na(activitywithNA$steps),]
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r echo=TRUE}
# Calculate the total number of steps for each day
stepsum <- aggregate(activity$steps, by=list(activity$date), FUN=sum, na.rm=TRUE)
# Assign column names
names(stepsum) <- c("date", "steps")
```
2. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
hist(stepsum$steps, breaks=53, xlab="Sum of Steps", col="red",  main="Histogram of Steps taken each day")
```
3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
# Calculate the mean of number of steps for each day
stepmean <- aggregate(activity$steps, by=list(activity$date), FUN=mean, na.rm=TRUE)
names(stepmean) <- c("date", "mean")
# Calculate the median of number of steps for each day
stepmedian <- aggregate(activity$steps, by=list(activity$date), FUN=median, na.rm=TRUE)
names(stepmedian) <- c("date", "median")
# Combine mean and median datasets
stepmeanandmedian <- cbind(stepmean,stepmedian$median)
# Update the column names
names(stepmeanandmedian) <- c("date", "mean", "median")
# display the result
stepmeanandmedian
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(activity$interval,activity$steps, type='l', xlab="", ylab="")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
step5min <- aggregate(activity$steps, by=list(activity$interval), FUN=sum, na.rm=TRUE)
names(step5min) <- c("interval",  "steps")
step5minordered <- step5min[order(step5min$steps),]
step5minordered[nrow(step5minordered), "interval"]

```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
#missing row count
nrow(activitywithNA) - nrow(activity)
```

## Are there differences in activity patterns between weekdays and weekends?
